---
title: "Trabajo_Bootstrap_Group2"
author: "Diego Alonso Gómez and ............."
date: "11/11/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bootstraping - Scripts in R and Python

For the bootstrap section you have to use the next equation:
log(inuidur1)~T4+ (female+black+othrace+factor(dep)+q2+q3+q4+q5+q6+agelt35+agegt54+durable+lusd+husd)

**No quadratic tem is required!!!**
Next, you have to compute the standard errors of 1,000 bootstrap estimates for the **T4, female** and **black**  variables.
Describe in detail each step you follow, and what lines of code you changed. 
Finally, present your results in a table. 

------

To remember the bootstrap concept, it is also necesary to know the meaning of "sampling with replacement": 
Sampling is called with replacement When a selected unit
randomly from the population is returned to the population and then a
second item at random. Whenever a unit is selected, the population
contains all the same units, so one unit can be selected
more than once.

-------------------------

## 1° Step: We open the dataset and subset it to compare treatment group 4 and the control group. 

```{r}

setwd("C:\\Users\\Diego\\Desktop\\Estadistica Aplicada\\Para trabajo grupal 11 de Noviembre")
library(boot)
Penn <- as.data.frame(read.table("../data/penn_jae.dat", header=T ))
n <- dim(Penn)[1]
p_1 <- dim(Penn)[2]
Penn<- subset(Penn, tg== 4| tg==0)
attach(Penn)
```


## 2° Setp: We define one of the reggresors, and then the function that will be included in the boot() function

```{r}
# Here we are defining one of the variable that would be included in our regression:
T4<- (tg==4)

# Here we have changed the form of the regression previously used
name <- log(inuidur1)~T4+ (female+black+othrace+factor(dep)+q2+q3+q4+q5+q6+agelt35+agegt54+durable+lusd+husd)  


# Here we the function that will define the form of our regression
boot.fn <- function(data, index) {
  coef(lm(log(inuidur1)~T4+ (female+black+othrace+factor(dep)+q2+q3+q4+q5+q6+agelt35+agegt54+durable+lusd+husd), data = data, subset = index)) }
#
```


```{r}
# Finally, here we define our bootstapping procces, what we only have changed is the name of our dataframe
example <- boot(Penn, boot.fn, 1000)
example
```

```{r}
# We do this to know which is the orden of our coefficents.
abc <- summary(lm(name, data = Penn))$coef
abc
```

```{r}
#Here we add a graphic: 
plot(example)
```

## Finally, we present our results in a table:

```{r}
table<- matrix(0, 4, 2)


table[1,1]<-abc[1,2]
table[2,1]<-abc[2,2]
table[3,1]<-abc[3,2]
table[4,1]<-abc[4,2]

table[1,2]<- 0.15384407
table[2,2]<- 0.03444049
table[3,2]<- 0.03551758
table[4,2]<- 0.05883169

colnames(table) <- c("std_with_linear_estimation", 'std_with_boot')
rownames(table)<- c('intercept','T4TRUE','female','black')

table
```
